CrypTool Script\footnote{ Background reading for the CrypTool program}\\ Mathematics and Cryptography

### 2003-06-06 script/en/chapters/ellipticcurves.tex
Änderungen Matthias

Translated freely, using the denotation of modern mathematics, this means: \\
No positive whole numbers $x, y$ and $z$ greater than zero exist such that $x^n +
y^n = z^n$ for $n>2$. I have found an amazing proof of this fact, but there is
too little space in the border [of the book] to write it down.

@@

Because the term {\em group} is used differently in everyday language than in
mathematics, we will, for reasons of completeness, begin by introducing the
essential statement of the formal definition of a group:
\begin{itemize}
   \item A group is a non-empty set $G$ and an operation $+.$ The set $G$ is
closed under the operation $+.$ Regardless of which two elements $a, b$ from $G$
are taken, performing the operation on them gives an element in $G$ (i.e.
$a+b=c$, and $c$ lies in $G$).
   \item For all elements $a, b$ and $c$ in $G$: $(a+b)+c = a+(b+c)$.
   \item There exists an element $e$ in $G$ that behaves neutrally with respect
to the operation $+$. That means that for all a in the set $G: ~a+e = e+a = a.$
   \item For each element $a$ in $G$ there exists a so-called inverse element $-a$
($-a$ also lies in $G$) such that: $a+(-a) = (-a)+a = e$.
\end{itemize}
If also $a+b = b+a$ for all $a, b$ in $G$ then we call the group an {\em
Abelian} group. An operation denoted as $+$ indicates an {\em additive} group;
if the operation is denoted as $\cdot$, we speak of a {\em multiplicative}
group.

@@

There are also {\em finite} groups. This means that these exists a set
$\mathcal{M}$ with a fixed number of elements and an operation $+$ such that the
above conditions are fulfilled. One example of this is any set ${\mathbb Z}_n$
where ${\mathbb Z}_n = \{0, 1, 2, 3, \cdots, n-1\}, n$ is a positive whole number
and the operation is addition mod $n$, i.e. $a$ and $b$ in ${\mathbb Z}_n$ are
subject to the operation $a+b \;{\rm mod~} n$.

\paragraph{Cyclic groups}
Cyclic groups\index{Group!cyclic} are those groups $G'$ that possess an element $g$
from which the group operation can be used to generate all other
elements in the group. This means that for each element $a$ in
$G'$ there exists a positive whole number $i$ such that if $g$ is
subject to the operation $i$ times (i.e. ``$g \cdot i$''),
$g+g+\cdots+g = a$ (additive group) or $g^i = g\cdot g \cdots g = a$
(multiplicative group). The element $g$ is the {\em generator} of
the cyclic group --- each element in $G´$ can be generated using
$g$ and the operation.

Now to the order of an element of the group: Let $a$ be in $G$. The smallest
positive whole number $r$ for which $a$ subject to the operation with itself $r$
times is the neutral element of the group $G´$ (i.e.: $r \cdot a = a+a+\cdots+a =
e$ bzw.\ $a^r = e$), is called the {\em order} of $a$.

@@

In mathematics, a field is understood to be a set $K$ with two operations
(denoted as $+$ and $\cdot$) which fulfils the following conditions:
\begin{itemize}
   \item The set $K$ forms an Abelian group together with the operation $+$
(addition), where $0$ is the neutral element of the operation $s$.
   \item The set $K$ (without the element 0) also forms an Abelian group
together with the operation $\cdot$ (multiplication).
   \item For all elements $a, b$ and $n$ in $K$, $n\cdot (a+b) = n \cdot a + n
\cdot b$ and $(a+b) \cdot n = a \cdot n + b \cdot n$.
\end{itemize}

There are {\em infinite} fields, i.e. the set on which the field is based
contains an infinite number of elements (e.g.: the field of real numbers). And
there are also finite fields, such as ${\mathbb Z}_p = \{0, 1, 2, 3, \cdots, p-1\}$
, where $p$ is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$ is a finite field.
\index{Field!Characteristic}
\paragraph{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with
respect to the multiplicative operation ``$\cdot$''. For positive
natural numbers $n$, let us understand $n_1$ to be $n_1 = 1 + 1 +
\cdots + 1$ ($n$ summands and $n_1$ is an element in $K$). If $n_1$
is then unequal to $0$ for all $n>0$, then we call $K$ a field
with characteristic zero. Otherwise, the characteristic of $K$ is
defined to be the smallest positive natural number $p$ for which
$p_1 = 0$ (note: $p$ is then a prime). Comment: The field of real
numbers has the characteristic $0$; the field ${\mathbb Z}_p$ has
the characteristic $p$.

@@

An elliptic curve is described by an equation. In order to keep it simple, we
restrict our explanation to elliptic curves over $${\mathbb Z}_p = \{0, 1, 2, 3,
\cdots, p-1\}$$ where $p$ is a prime greater than $3$. ${\mathbb Z}_p$ with
addition mod $p$ and multiplication mod $p$ is a finite field. However, we must
mention that elliptic curves can be defined over any (finite) field. In
particular, elliptic curves over fields with characteristic $2$ are extremely
interesting from a practical point of view because computers can be used to
represent the elements from these fields as bit strings. This leads to an
efficient implementation of the arithmetic in such fields, which means that a
computer can perform the operations of the field particularly quickly.

Because these points actually refer to the same thing, it is seldom necessary to distinguish between exact meanings.

An elliptic curve over ${\mathbb Z}_p$ is defined by an equation of the following form:
$$ y^2 \; ({\rm mod} \; p) = x^3 + ax + b ({\rm mod} \; p) $$
(thus: equality in the field ${\mathbb Z}_p$), where $a, b$ are in ${\mathbb Z}_p$ and $4a^3 + 27b^2$ mod $p$ is
not equal to zero. For fixed chosen numbers $a$ and $b$ in ${\mathbb Z}_p$, this equation has the pair of solutions
$$ {\bf E} = \left\{(x,y) \left| \begin{array}{c} x {\rm ~and~} y {\rm ~are~in~} {\mathbb Z}_p {\rm ~and~} \\
y^2  \equiv x^3 + ax + b \;({\rm mod~} p) {\rm ~and~} \\ 4a^3 + 27b^2 \not\equiv 0 \;
({\rm mod~} p)\end{array} \right. \right\}, $$ i.e. the set ${\bf E}$ consists of all pairs $x$ and $y$ that are a solution
(in ${\mathbb Z}_p$) of the above equation. It must be noted that the numbers $a, b$ and $p$ determine which pairs $(x,y)$
lie in the set ${\bf E}$. This means that $a, b$ and $p$ specify this set. The elements $(x,y)$ in ${\bf E}$ are called
points on the elliptic curve. In addition, ${\bf E}$ has one more element $O$ (the so-called point in infinity).
The set ${\bf E}$ is usually called an elliptic curve.

We can now define an operation\footnote{A animated addition of elliptic curve points can be
found at the web page of Certicom\index{Certicom}:
\href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\tt http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}.}
(also denoted as $+$, although it is
not the standard/usual addition of real numbers) on two elements in ${\bf E}$ such that the operation delivers an element
that also lies in ${\bf E}$. The set ${\bf E}$ is therefore closed under the operation $+$. We can show
that ${\bf E}$ is a group. The neutral element of the group ${\bf E}$ is the point in infinity $O$. Thus, for every two
points $(x_1,y_1)$ and $(x_2,y_2)$ on the elliptic curve ${\bf E}$, there exists a point $(x_3,y_3)$ on ${\bf E}$ such that
the operation $+$ complies with the following: $(x_1,y_1) + (x_2,y_2) = (x_3,y_3)$. Under certain circumstances, these points
may also be equal to the point in infinity. Thus, when we speak of a point $P$ on an elliptic curve ${\bf E}$,
we mean that $P = (x,y)$ and $(x,y)$ lies in the set ${\bf E}$. Any two points on an elliptic curve specified by $a, b$ and $p$
can therefore be added and the result is a point that also lies on the same elliptic curve.

% \newpage
\begin{figure}[htbp]
\subsubsection*{Adding of two points on an elliptic curve}
The following two figures show an elliptic curve in the affine plane and
shows how points on an elliptic curve are added. Note that the point in infinity $O$ cannot be represented in
the affine plane.
\begin{center}
\includegraphics[scale=1.08]{figures/ec-mult2}
\caption{Doubling the point P} \vspace{\floatsep} \vskip +20 pt
\includegraphics[scale=0.65]{figures/ec-add}
\caption{ Adding distinct points P and Q} % \footnotemark}
\end{center}
\end{figure}
% \addtocounter{footnote}{0}\footnotetext{The point $O$ cannot be represented in the affine plane.}
\enlargethispage{+20pt}

\newpage
We must note that ${\bf E}$ can have the following meanings:
\begin{itemize}
   \item the set ${\bf E}$ of solutions pairs $(x,y)$ for an equation including the point $O$
   \item the group ${\bf E}$ (with the operation ``addition of $(x_1,y_1)$ and $(x_2,y_2)$'')
   \item the elliptic curve ${\bf E}$ (which is actually the same as the group ${\bf E}$)
\end{itemize}
For cryptography, the important fact is that, for very large numbers, it appears to be extremely difficult to use a
given point $Q$ on an elliptic curve to determine which two points have to be added to obtain $Q$.

For large numbers $a, b$ and $p$ ($p$, for example, has a length of more than $160$ bits), the computer can easily add
the point $P$  $m$ times after another, i.e. to determine the point $P + P + \cdots + P = Q$ in an incredibly short space
of time (in a few fractions of a second) ($m$ summands $P$). Rather than $P + P + \cdots + P = Q$ ($m$ summands $P$)
we also write $mP = Q$. If we have a point $P$ and a point $Q$, which both lie on the same elliptic curve, no procedure
is known that enables us --- within an acceptable space of time --- to determine the number $m$ (assuming it actually exists)
for which $mP = Q$. This is referred to as the ``elliptic curve discrete logarithm problem'' (abbreviated to ECDLP)\index{ECDLP}.

We must note that not all elliptic curves are equally secure. This
means that we must choose the parameters $a$ and $b$ carefully
when defining a curve. For certain classes of elliptic curves, it
is possible to solve the ECDLP easier than in the general case.
Cryptographically unsuitable elliptic curves are called {\em abnormal}
curves (these are curves over ${\mathbb Z}_p$ for which the set
${\bf E}$ has precisely $p$ elements) and the {\em super-singular} curves
(the curves for which we can reduce the calculation of the ECDLP
to calculating the ``standard'' discrete logarithm in other finite
fields, i.e. simplify the calculation). There are therefore
cryptographically good and bad curves. However, for given
parameters $a$ and $b$ we can, with some difficulty, establish
whether or not the resulting elliptic curve is useful for
cryptographic purposes. The curves used in cryptography are
usually provided by experts. They ensure that the elliptic curves
they classify as secure satisfy the current security requirements.

For secure curves, the parameter $p$ determines how long it takes
to solve the ECDLP on this curve. The larger the parameter $p$,
the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it
clear why elliptic curves are so interesting for cryptography.
Because the parameter $p$ also determines the time required to
perform the signature/encryption procedure when using elliptic
curves in cryptography. The time taken to generate a pair of keys
also depends on $p$. Thus, small values (few bits) are desirable
here (in order to minimise the run times for the procedures);
however, the required security must still be maintained. For
example, with a length of $200$ bits for $p$, a {\em good}
elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the
current state of research). The reason for this is that the
quickest algorithms for solving the {\em elliptic curve discrete
logarithm} problem have an exponential run time --- unlike the
sub-exponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation
with elliptic curves). Therefore, the parameters for cryptographic
procedures based on the problem of {\em factorising whole numbers}
must be greater than the parameters for cryptographic procedures
based on the ECDLP problem.

\subsubsection{Digital signatures using elliptic curves}

The {\em elliptic curve discrete logarithm problem} (ECDLP) \index{ECDLP} forms the basis for
elliptic-curve cryptography. Various signature procedures are based on this.
What they have in common is how they use the public parameters and how these
parameters are used to generate secret and public keys:
\begin{itemize}
    \item The parameters of the elliptic curve E, i.e. a prime number $p$ that
determines over which field ${\mathbb Z}_p$ the elliptic curve ${\bf E}$ is
defined, as well as the two numbers $a$ and $b$ in ${\mathbb Z}_p$.
    \item A point $G=(x,y)$ that lies on the elliptic curve ${\bf E}$.
    \item A prime number $r < p$ for which $rG=O$ (i.e. $G$ added $r$ times
gives the neutral element of the group ${\bf E}$) and $r$ is a factor of $\#{\bf
E}$ (where $\# {\bf E}$ is the number of elements in the set ${\bf E}$). The
point $G$ therefore has the order $r$ and is the generator of a cyclic subgroup
of ${\bf E}$ with the order $r$.
    \item The number $k = \#{\bf E}/r$ ($k$ is called the {\em cofactor}).
\end{itemize}
The parameters $a, b, p, G, r$ and $k$ listed above are called
{\em domain parameters}.\index{Domain parameters} They determine on which elliptic curve ${\bf
E}$ and in which cyclic subgroup of ${\bf E}$ a signature
procedure has been ``used''.

The secret key $s$ of the signature generator is a (random\index{Random}) whole number $s$ in
the interval $[1, r-1]$. The public key of the signature generator is a point
$W=(x,y)$ on the elliptic curve ${\bf E}$. The public key $W$ and secret key $s$
are interrelated as follows: $W = sG$. This means that the domain parameters
(particularly $G$) and the secret key $s$ are used to calculate the public key
$W$ (by adding $G$ $s$ times on ${\bf E}$). The ECDLP is obviously used here: If
$W$ and $G$ (as well as the other domain parameters used) are known, then it is
difficult to use these to calculate $s$. (If the parameters are chosen
correctly, this currently appears to be practically impossible).

In order to verify a signature, the recipient of the signature must know the
following:
\begin{enumerate}
   \item The signature procedure used,
   \item The hash function used,
   \item The domain parameters used to generate the signature
   \item The public key $W$ of the signature generator.
\end{enumerate}

\subsubsection{Factorisation using elliptic curves}

>>>>

Translated freely, using the denotation of modern mathematics, this means: \\
No positive whole numbers $x, y$ and $z$ greater than zero exist such that $x^n +y^n = z^n$ for $n>2$. I have found an amazing proof of this fact, but there is
too little space in the border [of the book] to write it down.

@@

Because the term {\em group} is used differently in everyday language than in
mathematics, we will, for reasons of completeness, begin by introducing the
essential statement of the formal definition of a group:
\begin{itemize}
   \item A group is a non-empty set $G$ on which an operation ``$\cdot$''. The set $G$ is closed under this operation, which means that for any two elements $a, b$ taken from $G$, performing the operation on them gives an element in $G$, i.e. $ab=a\cdot b$ lies in $G$.
   \item For all elements $a, b$ and $c$ in $G$: $(ab)c = a(bc)$ (associative law).
   \item There exists an element $e$ in $G$ that behaves neutrally with respect
to the operation $\cdot$. That means that for all a in the set $G: ~ae = ea = a$.
   \item For each element $a$ in $G$ there exists a so-called inverse\footnote{The inverse is uniquely determined because if $x,y$ are inverse to $a$, then $x=xe=x(ay)=(xa)y=ey=y$.} element $a^{-1}$ in $G$ such that: $aa^{-1} = a^{-1}a = e$.
\end{itemize}
If also $ab = ba$ (commutative law) for all $a, b$ in $G$, then we call the group an {\em Abelian} group.

Since we may define different operations on the same set, we distinguish them by giving them different names (e.g. $+$ addition or $\cdot$ multiplication).

@@

For our purpose, so-called {\em finite} groups play an important role. This means that these exists a set
$\mathcal{M}$ with a fixed number of elements and an operation $+$ such that the
above conditions are fulfilled. One example of this is any set ${\mathbb Z}_n$
where ${\mathbb Z}_n = \{0, 1, 2, 3, \cdots, n-1\}, n$ is a positive whole number
and the operation is addition mod $n$, i.e. $a$ and $b$ in ${\mathbb Z}_n$ are
subject to the operation $a+b \;{\rm mod~} n$.

\paragraph{Cyclic groups}
Cyclic groups\index{Group!cyclic} are those groups $G'$ that possess an element $g$
from which the group operation can be used to generate all other
elements in the group. This means that for each element $a$ in
$G'$ there exists a positive whole number $i$ such that if $g$ is
subject to the operation $i$ times (i.e. ``$g\cdot i$''),
$g+g+\cdots+g = a$ (additive group) or $g^i = g\cdot g \cdots g = a$
(multiplicative group). The element $g$ is the {\em generator} of
the cyclic group --- each element in $G´$ can be generated using
$g$ and the operation.

\paragraph{Group order}
Now to the order of an element of the group: Let $a$ be in $G$. The smallest
positive whole number $r$ for which $a$ subject to the operation with itself $r$
times is the neutral element of the group $G´$ (i.e.: $r \cdot a = a+a+\cdots+a =
e$ respectively $a^r = e$), is called the {\em order} of $a$.

@@

In mathematics, one is often interested in sets on which at least two (group) operations are defined --- frequently called addition and multiplication. Most prominent are so called fields.

A field is understood to be a set $K$ with two operations
(denoted as $+$ and $\cdot$) which fulfils the following conditions:
\begin{itemize}
   \item The set $K$ forms an Abelian group together with the operation $+$
(addition), where $0$ is the neutral element of the operation $+$.
   \item The set $K\setminus\{ 0\}$ also forms an Abelian group
together with the operation $\cdot$ (multiplication).
   \item For all elements $a, b$ and $c$ in $K$, we have $c\cdot (a+b) = c \cdot a + c
\cdot b$ and $(a+b) \cdot c = a \cdot c + b \cdot c$ (distributive law).
\end{itemize}

Fields may contain an infinite number of elements (e.g. the field of real numbers). They are called {\em infinite} fields. In contrast we call a field
{\em finite}, if it contains only a finite number of elements (e.g. ${\mathbb Z}_p = \{0, 1, 2, 3, \cdots, p-1\}$
, where $p$ is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$).
\index{Field!Characteristic}
\paragraph{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with
respect to the multiplicative operation ``$\cdot$''. Then the characteristic of $K$ is said to be the order of $1$ with respect to the additive operation. This means that the characteristic of $K$ is the smallest positive integer $n$ such that
$$ \underbrace{1+1+\dots+1}_{\hbox{$n$ times}} =0 .
$$
If there is no such $n$, i.e. if $1+1+\dots+1\ne 0$ no matter how many $1$s we add, then we call $K$ a field
with characteristic $0$.

Thus, fields with characteristic $0$ are infinite since they contain the (pairwise distinct) elements $1$, $1+1$, $1+1+1$, \dots. On the other hand, fields with finite characteristic may by finite or infinite.

If the characteristic is finite, it has to be prime. This fact can easily be proved: Assume $n=pq$, $p,q<n$, is the characteristic of a field $K$. By definition of $n$, the elements $\bar p=\underbrace{1+1+\dots+1}_{\hbox{$p$ times}}$, $\bar q=\underbrace{1+1+\dots+1}_{\hbox{$q$ times}}$ of $K$ are not equal to $0$. Thus, there exist inverse elements $\bar p^{-1},\bar q^{-1}$ with respect to multiplication. It follows that $(\bar p\bar q)( $\bar p^{-1}\bar q^{-1})=1$, which contradicts the fact that $\bar p\bar q=\bar n=\underbrace{1+1+\dots+1}_{\hbox{$n$ times}}=0$ and, hence, $\underbrace{(\bar p\bar q)}_{=0}( $\bar p^{-1}\bar q^{-1})=0$.

Comment: The field of real
numbers has the characteristic $0$; the field ${\mathbb Z}_p$ has
the characteristic $p$. If $p$ is not prime, ${\mathbb Z}_p$ is not a field at all.

The most simple field is ${\mathbb Z}_2=\{ 0,1\}$. It contains only two elements, the neutral elements with respect to addition and multiplication. In particular, we have $0+0=0$, $0+1=1+0=1$, $1+1=0$, $1\cdot 1=1$, $0\cdot 0=0\cdot 1=1\cdot 0=0$.

\index{Finite!Fields}
\paragraph{Finite Fields}
As mentioned above, each finite field has a characteristic $p\ne 0$, where $p$ is a prime. On the other hand, given a prime $p$ there is a field which has exactly $p$ elements, that is ${\mathbb Z}_p$.

However, the number of elements of a field need not be prime in general. For example, it is not hard to construct a field with $4$ elements\footnote{%
The set $K=\{0,1,a,b\}$ fitted with the operation defined in the tabular below is a field:\\
$
\begin{array}{|c||c|c|c|c|} 
\hline 
+ & 0 & 1 & a & b \\
\hline \hline
0 & 0 & 1 & a & b \\
\hline 
1 & 1 & 0 & b & a \\
\hline 
a & a & b & 0 & 1 \\
\hline 
b & b & a & 1 & 0 \\
\hline 
\end{array} \qquad {\rm ~und~} \qquad
\begin{array}{|c||c|c|c|c|} 
\hline 
\cdot & 0 & 1 & a & b  \\
\hline \hline
0 & 0 & 0 & 0 & 0 \\ 
\hline 
1 & 0 & 1 & a & b \\ 
\hline 
a & 0 & a & b & 1 \\ 
\hline 
b & 0 & b & 1 & a \\
\hline 
\end{array} 
$  \\
}.

One can show that the order of any field is a prime power (i.e. the power of a prime number). On the other hand, we can construct a field with $p^n$ elements for any given prime $p$ and positive integer $n$. Since two fields that have the same number of elements can not be distinguished\footnote{If $K,K'$ are fields with $k=p^n$ elements, then there is a one-to-one map $\varphi:K\to K'$, that respects the arithmetic of the field. Such a map is called an isomorphy. Isomorphic fields mathematically behave in the same way so that dass it makes no sense to distinguish between them. For example, ${\mathbb Z}_2$ und $K'=\{ ZERO,ONE\}$ with zero-element $ZERO$ and one-element $ONE$ are isomorphic. We note that mathematical objects are only defined by their mathematical properties.}, we say that there is {\bf the field with $p^n$ elements} and denote it by $GF(p^n)$. Here $GF$ stands for {\it Galois Field} to commemorate the French Mathematician Galois.

The fields $GF(p)$ of prime order play a prominent role. They are called prime fields and often denoted by ${\mathbb Z}_p$\footnote{For prime fields additive as well as multiplicative group are cyclic. Furthermore, each field $GF(p^n)$ contains a subfield that is isomorphic to the prime field ${\mathbb Z}_p$.}.

@@

In cryptography elliptic curve are a useful tool. Such curves are described by some equation. A detailed analysis has shown that curves of the form\footnote{This curve is given by the zeros of a {\it polynomial}\index{Polynomial} $F$ of degree three in three variables. In general, expressions of the form
$P=\sum_{i_1,\dots,i_n\in\N_0} a_{i_1\dots i_n} x_1^{i_1}\dots x_n^{i_n}$ with coefficients $a_{i_1\dots i_n}\in K$ are called polynomials in $n$ variables $x_1,\dots,x_n$ with underlying field $K$, if ${\rm deg\,} P:=\max\{i_1+\dots +i_n: a_{i_1\dots i_n}\ne 0\}$ is finite, i.e. the sum has only finitely many non-zero terms (monomials).}
\begin{equation}
 F(x_1,x_2,x_3)=-x_1^3+x_2^2x_3+a_1x_1x_2x_3-a_2x_1^2x_3+a_3x_2x_3^2-a_4x_1x_3^2-a_6x_3^3=0,
\label{eccbasisgleichung}
\end{equation}
are especially useful. The variables $x_1,x_2,x_3$ and parameters $a_1,\dots,a_4,a_6$ are elements of a given field $K$, which has certain properties that are make it useful from the cryptographic point of view. The underlying field $K$ might be the well known field of real numbers or some finite field (see last section).
In order to obtain a cure that is useful for cryptography, the parameters have to be chosen in a way that the following conditions hold
$$ {\partial F\over\partial x_1}\ne 0, \quad {\partial F\over\partial x_2}\ne 0, \quad
{\partial F\over\partial x_3}\ne 0 .
$$
We identify points on the curve that can be derived from each over by multiplying each component with some scalar. This makes sense since $(x_1,x_2,x_3)$ solves (\ref{eccbasisgleichung}) if and only if $\alpha (x_1,x_2,x_3)$ ($\alpha\ne 0$) does. Formally, this means that we consider classes of equivalent points instead of single points, where points are called equivalent if one is the scalar multiple of the other one.
\\ If we put $x_3=0$ in the basic equation (\ref{eccbasisgleichung}), then this equation collapses to $-x_1^3=0$, leading to $x_1=0$. Thus, the equivalence class which includes the element $(0,1,0)$ is the only one that contains a point with $x_3=0$. For all points on the curve that are not equivalent to $(0,1,0)$, we may apply the following transformation
$$ K\times K\times (K\setminus\{0\})\ni (x_1,x_2,x_3) \mapsto (x,y):=\left( {x_1\over x_3}, {x_2\over x_3}\right) \in K\times K \, ,
$$
which reduces the number of variables to two instead of three. We note that the basic equation (\ref{eccbasisgleichung}) $F(x_1,x_2,x_3)=0$  was chosen in a way that this transformation leads to the famous so-called Weierstrass-Equation\footnote{Karl Weierstrass, 1815-1897, German mathematician, famous for his rigorous formal approach to mathematics.} holds
\begin{equation}
 y^2+a_1xy+a_3y = x^3+a_2x^2+a_4x+a_6 \, .
\label{ell}
\end{equation}
Since all but one point (i.e. equivalence class) of the elliptic curve can be described using equation (\ref{ell}), this equation is often called the elliptic equation, and its solutions written as
$$ {\bf E} = \left\{(x,y)\in K\times K \, |\, y^2+a_1xy+a_3y = x^3+a_2x^2+a_4x+a_6  \right\} \cup \{{\cal O} \}.
$$
Here, ${\cal O}$ represents the point $(0,1,0)$ that is loosely speaking mapped to infinity by the transformation (division by $x_3$) that reduces the three variables to two.

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.60]{figures/elliptic-curve}
\caption{Example of an elliptic curve with the real numbers as underlying field.\vspace{1ex}} 
\label{ExampleEllipticCurve}
\end{center}
\vskip -10 pt
\end{figure}

In contrast to figure \ref{ExampleEllipticCurve} only finite fields $K=GF(p^n)$ are used in elliptic curve cryptography. The reason is loosely speaking that in modern communication engineering data processing is always based on discrete data (simply because computers accept only discrete data).

For practical reasons, it turned out to be useful to take either $GF(p)$ with a large prime $p$ or $GF(2^n)$ with a (large) positive integer $n$. Using $GF(p)$ has the advantage of providing a relatively simple arithmetic; on the other hand $GF(2^n)$ allows a binary representation of each element that supports the way computers work. Other fields like, for example, $GF(7^n)$ do not have any of these advantages and are, thus, not considered, although there is no mathematical reason why they should not.

A coordinate transformation can result in a simpler version\footnote{Such a coordinate transformation is combination of a rotation and a dilatation of the coordinate system without changing the elliptic curve itself.} of the Weierstrass-equation. Depending whether $p>3$, different transformations are used, and we obtain
\begin{itemize}
\item in case of $GF(p)$, $p>3$, the elliptic curve equation of the form
\begin{equation}
 y^2 = x^3 + ax + b
\label{ellp}
\end{equation}
with $4a^3+27b^2\ne 0$
\item in case of $GF(2^n)$ the elliptic curve equation of the form 
\begin{equation}
 y^2+xy = x^3 + ax^2 + b
\label{ell2}
\end{equation}
with $b\ne 0$\footnote{The form (\ref{ellp}) is called the standard form of the Weierstrass-equation. If the characteristic of the field is $2$ or $3$, we obtain $4=0$ respectively $27=0$, which means that the condition on parameters $a,b$ collapse. Loosely speaking, this is the reason why the transformation to the standard form does not work in these cases.}.
\end{itemize}
This conditions on the parameters $a,b$ ensure that the elliptic equation can be used in the context of cryptography\footnote{Formally we call such curves non singular.}.

Let $|E|$ denote the number of elements of an elliptic curve $E$ given an underlying field $GF(k)$ (for practical reasons either $k=p$ with $p$ prim or $k=2^n$). Then Hasse's theorem\cite{Silverman1986} yields $| \, |E| - k-1\,| \le 2\cdot \sqrt{k}$. This Inequality is equivalent to $k+1 - 2\sqrt{k} < |E| < k+1+2\sqrt{k}$. In particular, this means that the number of elements of an elliptic curve is approximately $k$ (for large $k$).

% -----------------------------------------------------------------------------
\subsection{Operating on the elliptic curve}

In order to work with elliptic curves in practice, we define an operation (often written in an additive way $+$) on the set of points on the curve. If we have a curve over the field $GF(p)$, we define the commutative operation $+$ by
\begin{enumerate}
\item $P+{\cal O}={\cal O}+P=P$ for all $P\in E$,
\item for $P=(x,y)$ and $Q=(x,-y)$ we set $P+Q={\cal O}$,
\item for $P_1=(x_1,x_2),P_2=(x_2,y_2)\in E$ with $P_1,P_2\ne {\cal O}$ and $(x_2,y_2)\ne (x_1,-y_1)$ we set $P_3:=P_1+P_2$, $P_3=(x_3,y_3)$ defined by
$$ x_3:=-x_1-x_2+\lambda^2 \, , \qquad y_3:=-y_1+\lambda (x_1-x_3)
$$
with the auxiliary quotient
$$ \lambda:=\left\{ \begin{array}{cl} {y_1-y_2\over x_1-x_2} & {\rm if~} P_1\ne P_2, \\
                                     {3x_1^2+a\over 2y_1} & {\rm if~} P_1=P_2. \end{array} \right.
$$
\end{enumerate}
In particular, we obtain $-P=(x,-y)$ for $P=(x,y)\in E$.

If we deal with a curve over the field $GF(2^n)$, we define the operation $+$ in an analogous way by
\begin{enumerate}
\item $P+{\cal O}={\cal O}+P=P$ for all $P\in E$,
\item for $P=(x,y)$ and $Q=(x,x+y)$ we set $P+Q={\cal O}$,
\item for $P_1=(x_1,x_2),P_2=(x_2,y_2)\in E$ with $P_1,P_2\ne {\cal O}$ and $(x_2,y_2)\ne (x_1,x_1+y_1)$ we set $P_3:=P_1+P_2$, $P_3=(x_3,y_3)$ defined by
$$ x_3:=-x_1+x_2+\lambda+\lambda^2+a \, , \qquad y_3:=y_1+x_3+\lambda (x_1+x_3)
$$
with auxiliary quotient
$$ \lambda:=\left\{ \begin{array}{cl} {y_1+y_2\over x_1+x_2} & {\rm if~} P_1\ne P_2, \\
                                   x_1+{y_1\over x_1} & {\rm if~} P_1=P_2. \end{array}\right.
$$
\end{enumerate}
In particular, we obtain $-P=(x,-y)$ for $P=(x,y)\in E$.

(Note that $-(-P)=(x,x+(x+y))=(x,2x+y)=(x,y)$, since the underlying field has characteristic $2$.)\footnote{An animation of the addition of points on elliptic curves can be found on the Certicom\index{Certicom}-Homepage \\
\href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\texttt{http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}}}

% \newpage
\begin{figure}[htbp]
\subsubsection*{How to add points on an elliptic curve}
The following figures show how points on an elliptic curve over the field of real numbers are summed up using affine coordinates.
We note that the point infinity ${\cal O}$ cannot be shown in the affine plane.  
\begin{center}
\includegraphics[scale=1.08]{figures/ec-mult2}
\caption{Doubling of a point} 
\vspace{\floatsep}
\vskip +20 pt
\includegraphics[scale=0.65]{figures/ec-add}
\caption{Summing up two different points over the real number field} % \footnotemark }
\end{center}
\end{figure}
\enlargethispage{+20pt}
\newpage

% -----------------------------------------------------------------------------
\subsection{Security of elliptic-curve-cryptography: The ECDLP}

As mentioned above in section \ref{ECC-Crypto}, we only consider elliptic curves over the finite fields $GF(2^n)$ or $GF(p)$ (for a large prime $p$). This means that all parameters that describe the curve are taken from this underlying field. If $E$ is an elliptic curve over such a field and $P$ is a point on the curve $E$, then we can derive for all positive integers $m$
$$ mP := \underbrace{P+P+\dots+P}_{\hbox{$m$ times}} \, .
$$
Looking on this operation from the cryptographic point of view, it turns out to be very interesting by the following reason: On the one hand one needs only $\log m$ operations to calculate $mP$ --- one simply has to calculate $P$, $2P$, $2^2P$, $2^3P$, \dots, write $m$ in a binary form and finally add all these multilples $2^kP$ of $P$ with respect to the binary representation of $m$ --- on the other hand it seems to be very hard to find $m$ given $P$ and $Q=mP$ on $E$. Of course, we may simply calculate $P,2P,3P,4P,5P,\dots$ and compare each of them with $Q$. But this will take as much as $m$ operations.

Yet there is no algorithm known that efficiently derives $m$ given $P$ and $G$. The best algorithms known so far need about $\sqrt{q}$ operations where $q$ is the (largest) prime factor of $p-1$, in case the underlying field is $GF(p)$; here $m$ should be between $1$ and $q$ liegen so that one needs at most $\log q$ operations to calculate $mP$. However, the quotient ${\sqrt{q}\over\log q}$ tends to $+\infty$ very fast for large $q$.

If we choose the parameters sufficiently large (for example, let $p$ be prime and at least $160$ bits long), an computer will easily be able to calculate $mP$ (in less than a second). The {\it inverse problem} however, to derive $m$ from $mP$ and $P$, can (still) not be solved in reasonable time.

This problem is known as the \glqq over Elliptic Curve Discrete Logarithm Problem \grqq\ (for short ECDLP\index{ECDLP}).

\vskip +5 pt

In elliptic curve cryptography we formally look at points on the elliptic curve as elements of a group with point addition $+$ as operation. Furthermore, we use only elliptic curves that have a sufficiently large number of points. However, in special cases curves may be weak and not useful due to other reasons. For such special cases the ECDLP can be much easier to solve than in the general case. This means that one has to look carefully at the parameters when choosing an elliptic curve for cryptographic applications.

Not useful for cryptography are {\em a-normal} (that are curves over ${\mathbb Z}_p$,
for which the set ${\bf E}$ consists of exactly $p$ elements) and {\em supersingular} curves (that are curves, for which the ECDLP can be reduced to the ``normal'' discrete logarithms in another, smaller finite field). This means that there are cryptographically useful and non-useful elliptic curves. Given the parameters $a$ and $b$, it is possible to determine whether a curve is useful or not. In many publications one can find parameters that turned out to be useful for cryptography. The open (scientific) discussion guarantees that these results take into account latest research.
\vskip +5 pt
